---
title:  "p28443 common cells"
author: "Natalia Zajac"
output:
  html_document: 
    highlight: pygments
    theme: sand
    code_folding: hide
    toc: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---

Started on `r format(Sys.time(), "%Y-%m-%d %H:%M:%S")`

```{r setup, include=FALSE}
library(SummarizedExperiment)
library(plotly)
library(scater)
library(ezRun)
library(DT)
library(htmltools)
library(Matrix)
library(Seurat)
library(pheatmap)
library(DropletUtils)
library(cowplot)
library(tidyverse)
library(ggvenn)
library(patchwork)
library(UpSetR)
library(clusterProfiler)
library(org.Hs.eg.db)
library(BiocGenerics)
library(cowplot)
library(gridExtra)
library(ggpubr)

source("functions.R")
sample_data = ezRead.table("../2024-05-22-sampleInfo.tsv")

geneAnno <- ezRead.table("/srv/GT/reference/Homo_sapiens/GENCODE/GRCh38.p13/Annotation/Release_39-2021-12-09/Genes/SMRTLink11.1_annotation/genes_annotation_byGene.txt")
#Read in mtcontent from Hubert
mtcontent = read.csv("/srv/GT/analysis/zajacn/p28443/mito_content.csv")


illDirs <- file.path("/srv/gstore/projects", sample_data$CellRangerNoIntron, "filtered_feature_bc_matrix/")

seurat_IllDataList <- lapply(illDirs, function(crDir){
  cts_Illumina <- Read10X(crDir, gene.column = 1)
  featInfo <- ezRead.table(paste0(crDir, "/features.tsv.gz"), header = FALSE, row.names = NULL) # , col_names = FALSE)
  colnames(featInfo) <- c("gene_id", "gene_name", "type")
  featInfo <- featInfo[featInfo$gene_name %in% geneAnno[geneAnno$type == "protein_coding",]$gene_name, ]
  featInfo <- featInfo[!grepl("(?i)^MT-", featInfo$gene_name),]
  featInfo$isRiboprot <- grepl("(?i)^RPS|^RPL", featInfo$gene_name)
  geneInfoCr <- featInfo
  ctsGeneCr <- averageRows(cts_Illumina[featInfo$gene_id,], by = geneInfoCr$gene_name, func = sum) %>% as.matrix() %>% 
    as("dgCMatrix")
  seurat_object_Illumina =  CreateSeuratObject(ctsGeneCr)
  seurat_object_Illumina <- PercentageFeatureSet(seurat_object_Illumina, "(?i)^RPS|^RPL", col.name = "percent_ribo")
  colnames(seurat_object_Illumina) = sapply(str_split(colnames(seurat_object_Illumina), "-"), .subset, 1)
  samplename = sample_data[sample_data$CellRangerNoIntron == str_remove(str_remove(crDir, "/srv/gstore/projects/"), "/filtered_feature_bc_matrix/"),]$Name
  mt = mtcontent[mtcontent$Barcode %in% colnames(seurat_object_Illumina) & mtcontent$Sample == samplename,]
  seurat_object_Illumina$percent_mito = setNames(mt[, c(5)], mt[, c(3)])
  return(seurat_object_Illumina)
})


names(seurat_IllDataList) = sample_data$Name

sample_data$PigeonUnfiltered_allCells = list.dirs("/srv/gstore/projects/p28443/o31598_SMRTLink_IsoSeq/p28443_o31598_allsamples_allcells_pigeon/")[grepl("genes_seurat",list.dirs("/srv/gstore/projects/p28443/o31598_SMRTLink_IsoSeq/p28443_o31598_allsamples_allcells_pigeon/"))]


seurat_pbDataList <- lapply(sample_data$PigeonUnfiltered_allCells, function(pbDir){
  samplename = sample_data[sample_data$PigeonUnfiltered_allCells == pbDir,]$Name
  cts <- Read10X(pbDir, gene.column = 1)
  colnames(cts) = sub("-1", "", colnames(cts)) %>% DNAStringSet() %>% reverseComplement() %>% as.character()
  cts = cts[,colnames(seurat_IllDataList[[samplename]])]
  featInfo <- ezRead.table(paste0(pbDir, "/genes.tsv"), header = FALSE, row.names = NULL) # , col_names = FALSE)
  colnames(featInfo) <- c("gene_id", "gene_name")
  featInfo <- featInfo[featInfo$gene_name %in% geneAnno[geneAnno$type == "protein_coding",]$gene_name, ]
  featInfo <- featInfo[!grepl("(?i)^MT-", featInfo$gene_name),]
  featInfo$isRiboprot <- grepl("(?i)^RPS|^RPL", featInfo$gene_name)
  geneInfo <- featInfo
  ctsGene <- averageRows(cts[featInfo$gene_id,], by = geneInfo$gene_name, func = sum) %>% 
    as.matrix() %>% as("dgCMatrix")
  seurat_object = CreateSeuratObject(ctsGene)
  seurat_object <- PercentageFeatureSet(seurat_object, "(?i)^RPS|^RPL", col.name = "percent_ribo")
  mt = mtcontent[mtcontent$Barcode %in% colnames(seurat_object) & mtcontent$Sample == samplename,]
  seurat_object$percent_mito = setNames(mt[, c(4)], mt[, c(3)])
  return(seurat_object)
})
names(seurat_pbDataList) = sample_data$Name

pbDirs <- file.path(sample_data$PigeonUnfiltered, "genes_seurat/")
seurat_pbDataList_realcells <- lapply(pbDirs, function(pbDir){
  samplename = sample_data[sample_data$PigeonUnfiltered == str_remove(pbDir, "/genes_seurat/"),]$Name
  cts <- Read10X(pbDir, gene.column = 1)
  colnames(cts) = sub("-1", "", colnames(cts)) %>% DNAStringSet() %>% reverseComplement() %>% as.character()
  featInfo <- ezRead.table(paste0(pbDir, "/genes.tsv"), header = FALSE, row.names = NULL) # , col_names = FALSE)
  colnames(featInfo) <- c("gene_id", "gene_name")
  featInfo <- featInfo[featInfo$gene_name %in% geneAnno[geneAnno$type == "protein_coding",]$gene_name, ]
  featInfo <- featInfo[!grepl("(?i)^MT-", featInfo$gene_name),]
  featInfo$isRiboprot <- grepl("(?i)^RPS|^RPL", featInfo$gene_name)
  geneInfo <- featInfo
  ctsGene <- averageRows(cts[featInfo$gene_id,], by = geneInfo$gene_name, func = sum) %>% 
    as.matrix() %>% as("dgCMatrix")
  seurat_object = CreateSeuratObject(ctsGene)
  seurat_object <- PercentageFeatureSet(seurat_object, "(?i)^RPS|^RPL", col.name = "percent_ribo")
  mt = mtcontent[mtcontent$Barcode %in% colnames(seurat_object) & mtcontent$Sample == samplename,]
  seurat_object$percent_mito = setNames(mt[, c(4)], mt[, c(3)])
  return(seurat_object)
})
names(seurat_pbDataList_realcells) = sample_data$Name

```

# {.tabset}

## nFeature_RNA per cell

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
for (i in names(seurat_IllDataList)){
  print(merge(merge(data.frame("Illumina" = seurat_IllDataList[[i]]$nFeature_RNA), data.frame("PacBio" = seurat_pbDataList[[i]]$nFeature_RNA), by = "row.names"), data.frame("pacbio_real" = seurat_pbDataList_realcells[[i]]$nFeature_RNA), by.x = "Row.names", by.y = "row.names", all = TRUE) %>% mutate(real_cells = if_else(is.na(pacbio_real) == TRUE, "not_real_cell", "real_cell")) %>% ggplot(aes(Illumina,PacBio, color = real_cells)) + geom_point() + ggtitle(i))
}
```

## nCount_RNA per cell

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
for (i in names(seurat_IllDataList)){
  print(merge(merge(data.frame("Illumina" = seurat_IllDataList[[i]]$nCount_RNA), data.frame("PacBio" = seurat_pbDataList[[i]]$nCount_RNA), by = "row.names"), data.frame("pacbio_real" = seurat_pbDataList_realcells[[i]]$nCount_RNA), by.x = "Row.names", by.y = "row.names", all = TRUE) %>% mutate(real_cells = if_else(is.na(pacbio_real) == TRUE, "not_real_cell", "real_cell")) %>% ggplot(aes(Illumina,PacBio, color = real_cells)) + geom_point() + ggtitle(i))
}
```

## nCount_RNA per cell Zoom in

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
for (i in names(seurat_IllDataList)){
  print(merge(merge(data.frame("Illumina" = seurat_IllDataList[[i]]$nCount_RNA), data.frame("PacBio" = seurat_pbDataList[[i]]$nCount_RNA), by = "row.names"), data.frame("pacbio_real" = seurat_pbDataList_realcells[[i]]$nCount_RNA), by.x = "Row.names", by.y = "row.names", all = TRUE) %>% mutate(real_cells = if_else(is.na(pacbio_real) == TRUE, "not_real_cell", "real_cell")) %>% ggplot(aes(Illumina,PacBio, color = real_cells)) + geom_point() + xlim(0,30000) + ylim(0,30000) + ggtitle(i) )
}
```


## Empty droplet analysis {.tabset}

### Illumina {.tabset}

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}

library(DropletUtils)

illDirs_raw = str_replace(illDirs, "filtered", "raw")

illDirs_raw = lapply(illDirs_raw, function(x) {
  features = read.delim(paste0(x,"/features.tsv.gz"), header = F)
  mtx = read10xCounts(x)
  rownames(mtx) = features$V2
  features$isMito = grepl( "(?i)^MT-", features$V2)
  features$isRiboprot = grepl(  "(?i)^RPS|^RPL", features$V2)
  mtx = mtx[!features$isMito & !features$isRiboprot, ]
  return(mtx)
})

names(illDirs_raw) = sample_data$Name

pacbioDirs_raw = lapply(sample_data$PigeonUnfiltered_allCells, function(x) {
  features = read.delim(paste0(x,"/genes.tsv"), header = F)
  mtx = read10xCounts(x)
  features$isMito = grepl( "(?i)^MT-", features$V2)
  features$isRiboprot = grepl(  "(?i)^RPS|^RPL", features$V2)
  mtx = mtx[!features$isMito & !features$isRiboprot, ]
  return(mtx)
})

names(pacbioDirs_raw) = sample_data$Name

```


#### UMI count vs rank plot

Illumina

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
bcrank_all = NULL
for (i in names(illDirs_raw)){
  bcrank <- barcodeRanks(counts(illDirs_raw[[i]]))
  bcrank$Sample = i

# Only showing unique points for plotting speed.
  uniq <- !duplicated(bcrank$rank)
  bcrank$knee = metadata(bcrank)$knee
  bcrank$inflection = metadata(bcrank)$inflection
  bcrank_all = rbind(bcrank_all, bcrank[uniq,])
  plot(bcrank$rank[uniq], bcrank$total[uniq], log="xy",
      xlab="Rank", ylab="Total UMI count", cex.lab=1.2)
  
  abline(h=metadata(bcrank)$inflection, col="darkgreen", lty=2)
  abline(h=metadata(bcrank)$knee, col="dodgerblue", lty=2)
  
  legend("bottomleft", legend=c("Inflection", "Knee"), 
          col=c("darkgreen", "dodgerblue"), lty=2, cex=1.2)
}
```

All Samples together

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(bcrank_all, aes(rank ,total, color = factor(Sample, levels = sample_data$Name))) + 
  geom_point() + 
  scale_color_manual(values = myColors) + 
  scale_x_log10() + 
  scale_y_log10() + 
  geom_hline(yintercept = 52122, color = myColors[["Normal"]]) + 
  geom_hline(yintercept = 19140, color = myColors[["ccRCC_2"]]) + 
  geom_hline(yintercept = 34617, color = myColors[["ccRCC_3"]]) + 
  geom_hline(yintercept = 28813, color = myColors[["ccRCC_4"]]) + 
  geom_hline(yintercept = 201, color = myColors[["ccRCC_5"]]) + 
  geom_hline(yintercept = 5077, color = myColors[["Normal"]], linetype = "dashed") + 
  geom_hline(yintercept = 1021, color = myColors[["ccRCC_2"]], linetype = "dashed") + 
  geom_hline(yintercept = 7352, color = myColors[["ccRCC_3"]], linetype = "dashed") + 
  geom_hline(yintercept = 1217, color = myColors[["ccRCC_4"]], linetype = "dashed") + 
  geom_hline(yintercept = 111, color = myColors[["ccRCC_5"]], linetype = "dashed")

```

PacBio

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
for (i in names(pacbioDirs_raw)){
  bcrank <- barcodeRanks(counts(pacbioDirs_raw[[i]]))

# Only showing unique points for plotting speed.
  uniq <- !duplicated(bcrank$rank)
  plot(bcrank$rank[uniq], bcrank$total[uniq], log="xy",
      xlab="Rank", ylab="Total UMI count", cex.lab=1.2)
  
  abline(h=metadata(bcrank)$inflection, col="darkgreen", lty=2)
  abline(h=metadata(bcrank)$knee, col="dodgerblue", lty=2)
  
  legend("bottomleft", legend=c("Inflection", "Knee"), 
          col=c("darkgreen", "dodgerblue"), lty=2, cex=1.2)
}
```


Per sample highlighting real cells and combining PacBio and Illumina

```{r, fig.width=6, fig.height=6, echo=FALSE, message=FALSE, warning=FALSE}

for (i in names(illDirs_raw)){
  bcrank <- barcodeRanks(counts(illDirs_raw[[i]]))
  bcrank$Sample = i
  
  inflection = metadata(bcrank)$inflection
  knee = metadata(bcrank)$knee
  
  uniq <- !duplicated(bcrank$rank)
  bcrank = bcrank[uniq,]
  bcrank = data.frame(bcrank) %>% arrange(total)
  
  cells_in_both = length(seurat_pbDataList_realcells[[i]]$nCount_RNA)
  cells_in_ill = length(seurat_pbDataList[[i]]$nCount_RNA)
  total_cells = nrow(bcrank)
  first_nb = total_cells - cells_in_both + 1
  second_nb = nrow(bcrank) - cells_in_ill + 1
  
  bcrank$color = "Rest"
  bcrank[first_nb:total_cells,]$color = "Common"
  bcrank[second_nb:(first_nb-1),]$color = "Unique to ILL"
  bcrank$color = factor(bcrank$color, levels = c("Common","Unique to ILL", "Rest"))
  
  bcrank2 <- barcodeRanks(counts(pacbioDirs_raw[[i]]))
  bcrank2$Sample = i
  
  inflection = metadata(bcrank2)$inflection
  knee = metadata(bcrank2)$knee
  
  uniq <- !duplicated(bcrank2$rank)
  bcrank2 = bcrank2[uniq,]
  bcrank2 = data.frame(bcrank2) %>% arrange(total)

  total_cells = nrow(bcrank2)
  first_nb = total_cells - cells_in_both + 1
  second_nb = nrow(bcrank2) - cells_in_ill + 1
  
  bcrank2$color = "Rest"
  bcrank2[first_nb:total_cells,]$color = "Common"
  bcrank2$color = factor(bcrank2$color, levels = c("Common","Rest"))
  
  
  filename = paste0(i, "UMI_KneePLOT.pdf")
  pdf(file.path("Figures_2025_RevisionsNAR",filename), width = 7, height = 7)
  print(
    ggplot() + geom_point(data = bcrank, aes(rank, total), color = "#A6761D") + geom_point(data = bcrank2, aes(rank, total), color = "cyan") + scale_x_log10() + scale_y_log10() + theme_bw() + labs(x = "Barcodes", y = "UMI Count") + scale_color_manual(values = c("#440154FF", "#A6761D", "grey")) + theme(legend.position = "top", legend.title = element_blank(), axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title.x = element_text(size = 15), axis.title.y = element_text(size = 15)) + geom_hline(yintercept = head(bcrank2[bcrank2$color == "Common",], n = 1)$total ,color = "cyan") + geom_hline(yintercept = head(bcrank[bcrank$color == "Common",], n = 1)$total ,color = "#A6761D") + geom_hline(yintercept = head(bcrank[bcrank$color == "Unique to ILL",], n = 1)$total ,color = "#A6761D", linetype = "dashed") + ggtitle(i)) 
    # +
    #annotate("text", x=100000, y=head(bcrank[bcrank$color == "Unique to ILL",], n = 1)$total + 200, label="Illumina specific Cells") +
    #annotate("text", x=100000, y=head(bcrank[bcrank$color == "Common",], n = 1)$total + 10000, label="Common Cells") +
    #annotate("text", x=100000, y=head(bcrank2[bcrank2$color == "Common",], n = 1)$total + 4000, label="Common Cells"))
  dev.off()
}

```

### PacBio {.tabset}

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}

pacbioDirs_raw_filt = lapply(pacbioDirs_raw, function(x){
  set.seed(100)
  emptyStats <- emptyDrops(x, niters=1e5)
  x <- x[,which(emptyStats$FDR <= 0.001)]
})


```


####  Distribution of pvalues for empty droplets

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}

for (i in names(illDirs_raw)){
  set.seed(100)
  limit <- 100   
  all.out <- emptyDrops(counts(pacbioDirs_raw[[i]]), lower=limit, test.ambient=TRUE)
  hist(all.out$PValue[all.out$Total <= limit & all.out$Total > 0],
       xlab="P-value", main="", col="grey80") 
}

```

#### Number of filtered cells per sample

```{r, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}

data.frame(nonemtpy_cells = t(data.frame(lapply(pacbioDirs_raw_filt, function(x) length(x$Barcode))))) %>% rownames_to_column() %>% datatable()

```


## Empty Droplet analysis BLAZE

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}

blaze = NULL
for (i in list.files("/srv/gstore/projects/p28443/o31598_SMRTLink_IsoSeq/o31598_BLAZE/", pattern = "emtpy_bc_list.csv")){
  empty = read.csv(paste0("/srv/gstore/projects/p28443/o31598_SMRTLink_IsoSeq/o31598_BLAZE/", i), header = F)
  blaze[[i]] = empty
}

names(blaze) = sample_data$Name
```


```{r, fig.width=6, fig.height=6, echo=FALSE, message=FALSE, warning=FALSE}

for (i in sample_data$Name){
  filename = paste0(i, "BLAZE_VENN.pdf")
  pdf(file.path("Figures_2025_RevisionsNAR",filename), width = 7, height = 7)
  print(ggVennDiagram::ggVennDiagram(list("ILL (CellRanger)" = colnames(seurat_IllDataList[[i]]), "BLAZE EmptyDrops" = blaze[[i]]$V1, "PB (Iso-seq)" = colnames(seurat_pbDataList_realcells[[i]])), set_color = c("#A6761D","black","darkcyan"), set_size = 6, label_size = 11) + ggtitle(i) + scale_x_continuous(expand = expansion(mult = .2)) + scale_fill_distiller(palette = "RdBu") +
          theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 20))) 
  dev.off()
}

```


Simulated Barcode rank plot

```{r, fig.width=6, fig.height=6, echo=FALSE, message=FALSE, warning=FALSE}

library(ggplot2)
library(dplyr)

# Parameters
set.seed(42)
n_real_cells <- 3000  # Number of high-confidence cells
n_multiplets <- 3000  # Number of multiplets
n_background <- 1000000  # Number of background barcodes

# Generate UMI counts
umi_real <- rlnorm(n_real_cells, meanlog = 9, sdlog = 0.3)  # High UMI counts
umi_multiplets <- rlnorm(n_multiplets, meanlog = 7.5, sdlog = 0.4)  # Intermediate UMI counts
umi_background <- rlnorm(n_background, meanlog = 5.5, sdlog = 0.5)  # Low UMI counts

# Combine and rank
umi_counts <- c(umi_real, umi_multiplets, umi_background)
barcode_data <- data.frame(
    Rank = seq_along(umi_counts),
    UMI_Counts = sort(umi_counts, decreasing = TRUE)
)

# Add fake thresholds for annotations
primary_knee <- which.max(diff(diff(log10(barcode_data$UMI_Counts[1:n_real_cells + n_multiplets]))))
secondary_knee <- n_real_cells + n_multiplets + which.max(diff(diff(log10(barcode_data$UMI_Counts[n_real_cells + n_multiplets:length(umi_counts)]))))

# Plot the barcode rank plot
px = ggplot(barcode_data, aes(x = Rank, y = UMI_Counts)) +
    geom_line(color = "blue", size = 1)  +
    scale_x_log10() +
    scale_y_log10() +
    labs(
        title = "Schematic Single-Cell Barcode Rank Plot",
        x = "Barcode Rank",
        y = "UMI Counts"
    ) +
    theme_bw() +
    theme(
        text = element_text(size = 14),
        plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        panel.grid.minor = element_blank()
    )

pdf("Figures_2025_RevisionsNAR/Simulated_barcode_rank_plot.pdf", width = 7, height = 6)
px
dev.off()
```



Percentage of unique Illumina cells identified as empty drops by BLAZE:

Normal - 222/(222 + 168)
ccRCC2 - 51/(51 + 137)
ccRCC3 - 162/(162 + 116)
ccRCC4 - 208/(208 + 365)
ccRCC5 - 365/(365 + 402)

